\name{}: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-03-29 paper/main.tex
Fix typo in graph convolution biosnap data preparation

\subsection{Evaluation and metrics}
\label{sec:evaluation}
To assess each model, we compute a variety of common metrics for
binary classification. As the datasets are highly imbalanced, we use
the area under the receiver operating characteristic curve
(AUROC) on training, validation and testing split. % We
% calculated the true positive rate (TPR), false positive rate (FPR), and
% precision score. We will further compute true positives (TP), false
% positives (FP), false negatives (FN) and finally true negatives (TN).

@@

Note that drugs and proteins can be interchanged in this formulation, and we
refer to the different measures as protein-centric microAUC
($MicroAUC_p$) and a drug-centric microAUC ($MicroAUC_d$).

@@

	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}
		\hline
		(c) Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
		\hline
		Naive predictor&DP pairs&$0.85$& --&--\\
		DTINet&DP pairs&$0.91$&$0.74$&$0.67$\\
		DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ \\
		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$\\
		\name{}&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$\\
		\hline
		\hline
		(d) Approach&\multicolumn{2}{c|}{Original scheme}&\multicolumn{2}{c|}{Protein split}\\
		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$\\
		\hline
		Naive predictor&DP pairs&$0.79$&-- &--\\
		DeepDTI&Drugs&$0.88$&$0.76$&$0.70$\\
		DeepDTA&DP pairs&$0.88$&$0.77$&$0.69$\\
		DeepConv-DTI&DP pairs&$0.88$&$0.76$&$0.73$\\
		MolTrans&DP pairs&$0.90$&$0.77$&$0.74$\\
		\hline
	\end{tabular}

@@

We trained, validated and finally tested all considered models on the
STITCH dataset using a 5-fold cross-validation over a protein split;
we then selected the best-performing models (with respect to $MicroAUC_p$,
see Section \ref{sec:evaluation}) and evaluate them in a 5-fold
protein-split cross-validation on the Yamanishi benchmark dataset to
avoid validation overfitting and yield more realistic testing
results. To evaluate the influence of the different features
separately, and to determine whether they ``localize'' on the PPI
graph (and therefore can be exploited successfully by the graph neural
networks), we train and evaluate models with different types of
features, and with and without inclusion of the PPI graph,
separately. We comparing the molecular (MolPred) and phenotype-based
(OntoPred) prediction model, and a combination of both where we
concatenate both types of features.  Table
\ref{tab:Results}\todo{check -- not currently in the table} shows the
results of these experiments.

>>>>

\subsection{Evaluation and metrics}

\subsubsection{Splitting schemes}

As DTI prediction is dependent on both drugs and proteins, there are multiple ways of determining training, validation and testing sets of pairs to evaluate each model. For cross-validation, we can perform the split over DTI pairs, over drugs and over proteins, respectively, where only in the latter two schemes unseen entities are shown to the model in the validation and testing phase. Additionally, these schemes have diverging expressiveness with respect to DTI prediction, as different information is known and leaked across training and testing phase.

\subsubsection{Metrics}
\label{sec:evaluation}
To assess each model, we compute a variety of common metrics for
binary classification. As the datasets are highly imbalanced, we use
the area under the receiver operating characteristic curve
(AUROC) on training, validation and testing split. % We
% calculated the true positive rate (TPR), false positive rate (FPR), and
% precision score. We will further compute true positives (TP), false
% positives (FP), false negatives (FN) and finally true negatives (TN).

@@

Note that drugs and proteins can be interchanged in this formulation. We
refer to the different measures as protein-centric microAUC
($MicroAUC_p$) and a drug-centric microAUC ($MicroAUC_d$).

@@

	\begin{tabular}{|l|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|p{1cm}|}

		\hline
		&&\multicolumn{3}{c|}{(c) Yamanishi results}&\multicolumn{3}{c|}{(d) BioSnap results}\\
		Approach&Original&Original scheme&\multicolumn{2}{c|}{Protein split}&Original scheme&\multicolumn{2}{c|}{Protein split}\\

		&Splitting scheme&Macro AUC&Macro AUC&Micro $AUC_p$&Macro AUC&Macro AUC&Micro $AUC_p$\\
		\hline
		Naive predictor&DP pairs&$0.85$& --&--&$0.79$&-- &--\\
		DTINet&DP pairs&$0.91$&$0.74$&$0.67$ & --&--&--\\
		DTIGEMS+&DP pairs&$\mathbf{0.93}$& $0.72$& $0.68$ & --&--&--\\
		DTI-CDF&Proteins&$0.85$&$\mathbf{0.85}$&$0.79$& --&--&--\\
		\hline
		DeepDTI&Drugs& --&--&--&$0.88$&$0.76$&$0.70$\\
		DeepDTA&DP pairs& --&--&--&$0.88$&$0.77$&$0.69$\\
		DeepConv-DTI&DP pairs& --&--&--&$0.88$&$0.76$&$0.73$\\
		MolTrans&DP pairs& --&--&--&$0.90$&$0.77$&$0.74$\\
		\name&Proteins&$0.84$&$0.84$&$\mathbf{0.84}$& & & \\
		\hline
	\end{tabular}\medskip

@@

We trained, validated and finally tested all considered models on the
STITCH dataset using a 5-fold cross-validation over a protein split;
we then selected the best-performing models (with respect to $MicroAUC_p$,
see Section \ref{sec:evaluation}) and evaluate them in a 5-fold
protein-split cross-validation on the Yamanishi benchmark dataset to
avoid validation overfitting and yield more realistic testing
results. To evaluate the influence of the different features
separately, and to determine whether they ``localize'' on the PPI
graph (and therefore can be exploited successfully by the graph neural
networks), we train and evaluate models with different types of
features, and with and without inclusion of the PPI graph,
separately. We comparing the molecular (MolPred) and phenotype-based
(OntoPred) prediction model, and a combination of both where we
concatenate both types of features.  Table
\ref{tab:Results} shows the
results of these experiments.
